The Ky Fan Norms and Beyond: Dual Norms and Combinations for Matrix Optimization

### 2025-11-26 conf_article/neon.tex
Added arrows
Fixed accuracies

\author{Alexey Kravatskiy \\
Moscow Institute of Physics and Technology (MIPT) \\
\texttt{kravtskii.aiu@phystech.edu} \\
\And
Ivan Kozyrev \\
Moscow Institute of Physics and Technology (MIPT), \\
Marchuk Institute of Numerical Mathematics \\
\texttt{kozyrev.in@phystech.edu} \\
\And
Nikolai Kozlov \\
Moscow Institute of Physics and Technology (MIPT) \\
\texttt{kozlov.na@phystech.edu} \\
\And
Alexander Vinogradov \\
Moscow Institute of Physics and Technology (MIPT) \\
\texttt{vinogradov.am@phystech.edu} \\
\And
Daniil Merkulov \\
Moscow Institute of Physics and Technology (MIPT), Skoltech, HSE, AI4Science \\
\texttt{daniil.merkulov@phystech.edu} \\
\And
Ivan Oseledets \\
AIRI, Skoltech \\
\texttt{i.oseledets@skoltech.ru}
}

@@

Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible. On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.04 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance. On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279. Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon. These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above. Moreover, the tools from ~\ref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that does not have to be modifications of Muon.

@@

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.8\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figs/simple_lls/loss_vs_iteration_long_run.pdf}
        \caption{The loss}
        \label{fig:lls_loss}
    \end{subfigure}
    \vspace{1em} % optional: add vertical spacing between subfigures
    \begin{subfigure}[t]{0.8\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_long_run.pdf}
        \caption{The Frobenius norm of the gradient}
        \label{fig:lls_fro_grad_norm}
    \end{subfigure}
    \caption{Linear least squares problem for a 500x500 matrix}
    \label{fig:lls}
\end{figure}

@@

Then we tune F-Muon with $\alpha = 0.5$. Tuned parameters are {\tt lr=0.42(1 - step/total\_steps), momentum=0.63, nesterov = True}. Tuned F-Muon reaches $94.04 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance. Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\cref{fig:fmuon_alphas}). We notice that even when $\alpha=0.1$, the accuracy is much higher than in the case of vanilla NSGD.

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|c|c|c|c|}
	  \hline
	  Method & lr  & momentum   & val\_accuracy, \% \\
	  \hline
	  NSGD & ? & ? & ? \\
	  \hdashline
	  SignSGD & ? $\cdot$ ? & ? & ? \\
	  \hdashline
	  Muon & 0.24 & 0.63 & $94.02 \pm 0.10$ \\
	  \hdashline
	  F-Muon & 0.42 & 0.63 & $94.04 \pm 0.13$\\
	  \hdashline
	  S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
	  \hline
	\end{tabular}
  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult}=0.003}
  \label{tbl:cifar_lrs}
  \end{table}

>>>>

\author{Alexey Kravatskiy \\
Moscow Institute of Physics and Technology (MIPT) \\
\texttt{kravtskii.aiu@phystech.edu} \\
\And
Ivan Kozyrev \\
Moscow Institute of Physics and Technology (MIPT), \\
Marchuk Institute of Numerical Mathematics \\
\texttt{kozyrev.in@phystech.edu} \\
\And
Nikolai Kozlov \\
Moscow Institute of Physics and Technology (MIPT) \\
\texttt{kozlov.na@phystech.edu} \\
\And
Alexander Vinogradov \\
Moscow Institute of Physics and Technology (MIPT) \\
\texttt{vinogradov.am@phystech.edu} \\
\And
Daniil Merkulov \\
Moscow Institute of Physics and Technology (MIPT), Skoltech, HSE, AI4Science \\
\texttt{daniil.merkulov@phystech.edu} \\
\And
Ivan Oseledets \\
AIRI, Skoltech, \\
Marchuk Institute of Numerical Mathematics \\
\texttt{i.oseledets@skoltech.ru}
}

@@

Most notably, our experiments on real-world tasks demonstrate that the choice of underlying matrix norm is remarkably flexible. On CIFAR-10 airbench, properly-tuned F-Muon and S-Muon achieve $94.02 \pm 0.13\%$ and $94.03 \pm 0.13\%$ accuracy, matching Muon's $94.01 \pm 0.13\%$ performance. On NanoGPT pre-training, F-Muon achieves 3.281 cross-entropy loss, while fully-tuned Muon achieves 3.279. Finally, S-Muon matches Muon on fine-tuning of NanoGPT on Tiny Stories, while F-Muon is far more resistant to the learning rate choice than Muon. These results show that Muon-like algorithms can maintain competitive performance even when the underlying norm constraint is significantly modified, answering affirmatively the central question posed above. Moreover, the tools from ~\ref{sec:conic_combo} give the researchers an unheard-of flexibility in designing algorithms that does not have to be modifications of Muon.

@@

\begin{figure}[h!]
    \centering
    \begin{subfigure}[t]{0.8\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figs/simple_lls/loss_long_run_16_increase.pdf}
        \caption{The loss}
        \label{fig:lls_loss}
    \end{subfigure}
    \vspace{1em} % optional: add vertical spacing between subfigures
    \begin{subfigure}[t]{0.8\linewidth}
        \centering
        \includegraphics[width=\linewidth]{figs/simple_lls/gradient_frobenius_norm_vs_iteration_long_run.pdf}
        \caption{The Frobenius norm of the gradient}
        \label{fig:lls_fro_grad_norm}
    \end{subfigure}
    \caption{Linear least squares problem for a 500x500 matrix}
    \label{fig:lls}
\end{figure}

@@

Then we tune F-Muon with $\alpha = 0.5$. Tuned parameters are {\tt lr=0.4(1 - step/total\_steps), momentum=0.6, nesterov = True}. Tuned F-Muon reaches $94.02 \pm 0.13\%$ after 8 epochs (averaged by 200 iterations), matching Muon's accuracy and variance. Once again we measure \texttt{val\_accuracy}($\alpha$) curve (\cref{fig:fmuon_alphas}). We notice that even when $\alpha=0.1$, the accuracy is much higher than in the case of vanilla NSGD.

@@

\begin{table}[ht]
	\centering
	\begin{tabular}{|c|c|c|c|}
	  \hline
	  Method & lr  & momentum   & val\_accuracy, \% \\
	  \hline
	  NSGD & ? & ? & ? \\
	  \hdashline
	  SignSGD & ? $\cdot$ ? & ? & ? \\
	  \hdashline
	  Muon & 0.24 & 0.6 & $94.01 \pm 0.10$ \\
	  \hdashline
	  F-Muon & 0.40 & 0.6 & $94.02 \pm 0.13$\\
	  \hdashline
	  S-Muon & 0.42 & 0.63 & $94.03 \pm 0.13$ \\
	  \hline
	\end{tabular}
  \caption{Parameters for CIFAR-airbench. \texttt{sign\_lr\_mult}=0.003}
  \label{tbl:cifar_lrs}
  \end{table}
